Uncertainty Quantification in Wind Power Forecasting

### 2020-03-28 LaTeX_Code/presentation/Palacio_WindPower.tex
I cleaned a bit the directories

\title{On the Uncertainty of Wind Power Generation}

@@

\begin{document}
\setbeamercolor{background canvas}{bg=blue!10}
\date{}

\begin{frame}
\titlepage
\center
  \includegraphics[width=40mm,scale=1]{KAUST_Logo.pdf}\quad 
  \includegraphics[width=60mm,scale=1]{rwth_UQ.png}\quad 
  \includegraphics[width=40mm,scale=1]{UParis_13.png} \quad
\end{frame}

@@

\begin{frame}\frametitle{Status quo }
  Wind power forecasts can be generally categorized as follows:
  \begin{itemize}
    \item physical models
    \item statistical methods
    \item artificial intelligence methods
    \item spatial correlation methods
    \item other hybrid approaches
  \end{itemize}
 The output of such methods is usually a \textcolor{red}{deterministic forecast}. Occasionally probabilistic forecasts are produced through uncertainty propagation in the data, parameters or through forecast ensembles. However, little has been done in terms of producing \textcolor{red}{data driven probabilistic forecasts} based on the real-world performance of forecasting models.
\end{frame}

\begin{frame}\frametitle{Data}
This is a year long data set from Uruguay based on \textbf{1000 72-hour long paths of observations every 10 min \textcolor{red}{ ($\sim$ half a million data points)}} recorded in 2018.
\begin{figure}
  \includegraphics[width=70mm,scale=1]{plots/data_1516064400.pdf}
  \includegraphics[width=70mm,scale=1]{plots/data_1516323600.pdf}\\
  \includegraphics[width=70mm,scale=1]{plots/data_1517619600.pdf}
  \includegraphics[width=70mm,scale=1]{plots/data_1529456400.pdf}
  % \caption{title}
\end{figure}
% Do you have more data? send it our way.
\end{frame}

@@

\begin{frame}\frametitle{Model}
Our goals are to
\begin{itemize}
  \item Generate a probabilistic forecast centered around a given deterministic forecast, i.e. unbiased with respect to the deterministic forecast.
  \item Capture the dynamics and correlation structure.
  \item Capture the skew nature of forecast errors.
  \item Be forecasting technology agnostic. Thus, compatible with future forecasting technology.
  \item Learn from historical power production data.
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Model}
We propose to model wind power forecasts errors using \textcolor{red}{parametric stochastic differential equations (SDEs)} whose solution defines a stochastic process. This resultant stochastic process describes the time evolution dynamics of wind power forecast errors.
\begin{equation}
\begin{split}
dX_t &= a(X_t; p_t,\bm{\theta}) dt + b (X_t;p_t, \bm{\theta} ) dW_t \quad t > 0 \\
X_0 & = x_0
\end{split}
\label{main}
\end{equation}

\begin{itemize}
\item $a(\cdot; \bm{\theta}):[0,1] \to \R $  a drift function.
\item $b (\cdot; \bm{\theta} ):[0,1] \to \R$  a  diffusion function.
\item $\bm{\theta}$: a vector of parameters.
\item $p_t$ time-dependent scalar value.
\item $W_t$: Standard Wiener random process in $\R$.
\end{itemize}

@@

\begin{frame}\frametitle{Model}
\textcolor{blue}{Answer:}
Let $\bm{\theta} = (\textcolor{red}{\theta_0}, \textcolor{red}{\alpha})$
\begin{enumerate}
  \item We want the process to follow the wind forecast, thus we choose a drift term that is mean reverting and tracks the derivative of the deterministic forecast $p_t$, which is an input to our model.
  \begin{equation}
    a(x; \bm{\theta})=  \dot{p_t} - \theta_t(x - p_t)
  \end{equation}
where $\theta_t$ is a time-dependent parameter that controls the speed of reversion.
  \item We want a diffusion term that vanishes at the boundaries to prevent the process from escaping the region $[0,1]$.
  \begin{equation}
    b (x; \bm{\theta} )= \sqrt{2 \theta_t \textcolor{red}{\textcolor{red}{\alpha}} x (1-x)}
  \end{equation}
  where $\textcolor{red}{\alpha}$ is a constant parameter that controls the path variability.

To further ensure that the process does not escape the region $[0,1]$, the mean reversion parameter has to be selected according to the following rule,
\begin{equation}
\theta_t = \max \left( \textcolor{red}{\theta_0} \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right ) \label{theta_t}
\end{equation}
\end{enumerate}
\end{frame}

@@

\begin{equation}
\mathcal{L}(\bm{\theta};V) =\prod\limits_{j=1}^M \prod\limits_{i=1}^N \rho ( {V_{j,i+1}|V_{j,i}}, \bm{\theta})  \rho (V_{j,0})
\label{likelihood}
\end{equation}

@@

\begin{equation}
\begin{split}
\frac{ \partial f }{\partial t } & (y ,t | x , s, \theta_t, \textcolor{red}{\alpha} )= - \frac{\partial}{ \partial y} ( a( y;\dot{p}_t , p_t, \theta_t ) f( y ,t | x , s, \theta_t , \textcolor{red}{\alpha} ) ) \\
& + \frac{1}{2} \frac{\partial^2}{ \partial y^2} ( b(y;\theta_t, \textcolor{red}{\alpha}  )  f( y ,t | x , s, \theta_t , \textcolor{red}{\alpha} )  ) \quad  t < s\\
\end{split}
\end{equation}
This is a parametric PDE which is \textcolor{red}{computationally expensive to solve} and optimize for every transition.
\end{frame}

\begin{frame}\frametitle{Moment Matching}
Instead of solving for exact transition densities by the Fokker-Planck, we propose a \textcolor{red}{proxy transition density}. We match the moments of our SDE model with that of the proxy density. Using Ito, we arrive at the following iterative ODEs.

\begin{equation}
\begin{split}
\frac{d \E[ V^k_t]}{dt} = - k \theta_t \E [ V^k_t] + \frac{k(k-1)}{2} \E [ V^{k-2}_t  b(V^k_t;\theta_t, \textcolor{red}{\alpha})]
\end{split}
\end{equation}
For $t\in [t_{n-1}, t]$, the first two moments are given by
\begin{equation}
\begin{split}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t \\
\frac{d m_2 (t)}{dt} &=  -2 m_2(t)\theta(1+\textcolor{red}{\alpha}) + 2\textcolor{red}{\alpha}\theta_t m_1(t)(1-2p_t) + 2 \textcolor{red}{\alpha}\theta_t p_t (1-p_t)\\
\end{split}
\end{equation}
with initial conditions, $m_1(t_{n-1})= v_{n-1}$ and $m_2(t_{n-1})= v_{n-1}^2$ where $t_{n-1}$ is the time of the previous obseravtion and $v_{n-1}$ is its value.

@@

Execute the following until an accuracy threshold is met:
\begin{enumerate}
\item \textcolor{red}{initialize}.
\item \textcolor{red}{optimize} the log-likelihood function.
\begin{enumerate}
\item For every evaluation of the log-liklihood function:
\begin{enumerate}
  \item \textcolor{red}{Sample} a random mini-batch of transitions and their associated forecast and parameters.
  \item \textcolor{red}{Solve} the ODE system for every transition to obtain the moments.
  \item \textcolor{red}{Match} the resulting moments with the parameters of the chosen proxy distribution for every transition.
\end{enumerate}
\end{enumerate}
\item go to step 1 (i.e. \textcolor{red}{re-initialize} the optimization with the most recent result ).
\end{enumerate}

@@

\begin{frame}\frametitle{Future Wind Power Simulation}
    \begin{figure}
      \includegraphics[width=75mm,scale=1]{simulated/24hr/1099.pdf}
      \includegraphics[width=75mm,scale=1]{simulated/24hr/1178.pdf}
      \caption{We simulate five possible future wind power production paths using the obtain optimal parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha} )=(12,0.3)$ }
    \end{figure}
\end{frame}

\begin{frame}\frametitle{Confidence Bands}
    \begin{figure}
      \includegraphics[width=75mm,scale=1]{confidence_intervals/24hr/31.pdf}
      \includegraphics[width=75mm,scale=1]{confidence_intervals/24hr/820.pdf}
       \caption{We obtain confidence intervals for future wind power production using the obtain optimal parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha} )=(12,0.3)$. Actual production plotted in retrospect. }
    \end{figure}
\end{frame}

\begin{frame}\frametitle{State-Independent model}
The model we have demonstrated previously is state-dependent, that is the diffusion of the SDE depends on the state.
\begin{equation}
\begin{split}
dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \textcolor{red}{\alpha} (V_t +p_t ) (1-V_t-p_t)} \  dW_t  \\ %\quad t > 0
V_0 & = v_0\\
\end{split}\label{VtSDE}
\end{equation}

Why are we interested in a state-independent model? Because it's more tractable.\\
We apply a \textbf{Lamperti transform} to obtain the following state-independent system,

\begin{equation}
  \begin{split}
    dZ_t&= \frac{- \theta_t (1+ \sin(Z_t) - 2p_t) + \textcolor{red}{\alpha} \theta_t \sin (Z_t)   }{\cos (Z_t)} + \sqrt{2 \textcolor{red}{\alpha} \theta_t} dW_t \\
    Z_0&=z_0
  \end{split}
\end{equation}
where $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.
\end{frame}

\begin{frame}\frametitle{Data Skewness after Lamperti transformation}
As stated before, the state-independent SDE follows a Lamperti transformed process $Z_t$ give by $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.
\begin{figure}
  \includegraphics[width=50mm,scale=1]{plots/hist_low_lamperti.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_mid_lamperti.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_high_lamperti.pdf}
  \caption{We observe that skewness has been greatly reduced after the Lamperti transformation. This motivates us to use a \textcolor{red}{Gaussian transition} density as a proxy density.}
\end{figure}
\end{frame}

\begin{frame}\frametitle{State-Independent model}
Similarly, we try to obtain a system of ODEs to determine the centered moments of the Lamperti transformed process $V_t$. Due to the non-linearity in the drift, we can only approximate the centered moments by the following ODEs,
\begin{equation}
\begin{split}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t (1-\textcolor{red}{\alpha}) - \theta (1-2 p_t) \\
\frac{d v(t)}{dt} &=  2 v(t) \theta_t (2p_t - 1 ) \tan(x) \sec(x) + \theta_t (\textcolor{red}{\alpha} - 1) \sec^2(x)  + 2 \theta_t \textcolor{red}{\alpha}\\
\end{split}
\end{equation}
These are not exact moment ODEs, however they are accurate enough for small time intervals.
\end{frame}

\begin{frame}\frametitle{Inference Results}
\begin{figure}
  \includegraphics[width=70mm,scale=1]{plots/alpha_conv_beta.pdf}
  \includegraphics[width=70mm,scale=1]{plots/theta_conv_beta.pdf}
    \caption{We have self-convergence of our algorithm in the Lamperti space at a rate that matches the convergence rate of Monte Carlo.}
\end{figure}
\end{frame}

\begin{frame}\frametitle{Model comparison}
\begin{table}[]
\centering
\begin{tabular}{|c|c|}
\hline
Model   &  parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha})$    \\ \hline
state-dependent &   (12,0.3)    \\ \hline
state-independent &   (14,0.29)     \\ \hline
\end{tabular}
\caption{We compare the parameters obtain in both the original and Lamperti space. }
\label{tab:model_comparison}
\end{table}
\end{frame}

\begin{frame}\frametitle{Concluding remarks}
We were able to:
\begin{itemize}
  \item simulate future wind power production based on real data.
  \item obtain an analytical description of the uncertainty of wind power forecasts in the form of an SDE.
  \item develop a forecasting technology agnostic method.
  \item capture skewness of the error process and its correlation structure.
\end{itemize}
\end{frame}

@@

\section*{References}
\begin{flushleft}
\begin{itemize}
\item Alhaddad, W. ,  Kebaier, A. , \& Tempone, R. (2019). Stochastic Wind Power Forecasting. Coming soon to  arXiv.
\item  M\o ller, J. K., Zugno, M., \& Madsen, H. (2016). Probabilistic Forecasts of Wind Power Generation by Stochastic Differential Equation Models. Journal of Forecasting, 35(3), 189-205. doi:10.1002/for.2367 
\item  Elkantassi, S., Kalligiannaki, E., \& Tempone, R. (2017). Inference And Sensitivity In Stochastic Wind Power Forecast Models. Proceedings of the 2nd International Conference on Uncertainty Quantification in Computational Sciences and Engineering (UNCECOMP 2017). doi:10.7712/120217.5377.16899
\item Giebel, G., Brownsword, R., Kariniotakis, G., Denhard, M., \& Draxl, C. (2011). The State-Of-The-Art in Short- Term Prediction of Wind Power: A Literature Overview, 2nd edition. ANEMOS.plus. https://doi.org/10.11581/DTU:00000017 
\item Chang, W.-Y. (2014) A Literature Review of Wind Forecasting Methods. Journal of Power and Energy Engineering, 2, 161-168. http://dx.doi.org/10.4236/jpee.2014.24023
\end{itemize}
\end{flushleft}

>>>>

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\setbeamercolor{footline}{fg=blue}
\setbeamerfont{footline}{series=\bfseries}

\title{Uncertainty Quantification in Wind Power Forecasting}

@@

\begin{document}
\setbeamercolor{background canvas}{bg=blue!10}
\date{October 4th, 2019}

\begin{frame}
\center
  \includegraphics[width=30mm,scale=1]{KAUST_Logo.pdf}\quad 
  \includegraphics[width=40mm,scale=1]{rwth_UQ.png}\quad 
  \includegraphics[width=30mm,scale=1]{UParis_13.png} \quad
 \vspace*{0.2in}
\titlepage
\end{frame}

@@

\begin{frame}\frametitle{Current State of Affairs}
  Wind power forecasts can be generally categorized as follows:
  \begin{itemize}
    \item physical models
    \item statistical methods
    \item artificial intelligence methods
    \item other hybrid approaches
  \end{itemize}
 The output of such methods is usually a \textcolor{red}{deterministic forecast}. \\
 \bigskip 
 Occasionally probabilistic forecasts are produced through uncertainty propagation or through forecast ensembles. \\
  \bigskip 
 However, there is a lacking in  \textcolor{red}{data driven stochastic forecasts} based on the real-world performance of forecasting models.
\end{frame}

\begin{frame}\frametitle{Data}
This is a year long data set from Uruguay based on \textbf{1000 72-hour long paths with observations recorded  every 10 min \textcolor{red}{ ($\sim$ half a million data points)}} recorded in 2018.
\begin{figure}
  \includegraphics[width=70mm,scale=1]{plots/data_1516064400.pdf}
  \includegraphics[width=70mm,scale=1]{plots/data_1516323600.pdf}\\
  \includegraphics[width=70mm,scale=1]{plots/data_1517619600.pdf}
  \includegraphics[width=70mm,scale=1]{plots/data_1529456400.pdf}
  % \caption{title}
\end{figure}
% Do you have more data? send it our way.
\end{frame}

@@

\begin{frame}\frametitle{Model}

Our goals are to produce a stochastic forecast of wind power production forecasting errors while:
\begin{itemize}
  \item Capturing the dynamics of the forecast error process.
  \item Capturing the skew nature of forecast errors.
  \item Being forecasting-technology agnostic. Thus, compatible with past and future forecasting-technology.
  \item Learning from historical power production data.
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Model}
We propose to model wind power forecasts errors using \textcolor{red}{parametric stochastic differential equations (SDEs)} whose solution defines a stochastic process. This resultant stochastic process describes the time evolution dynamics of wind power forecasting errors.
\begin{equation}
\begin{split}
dX_t &= a(X_t; p_t, \dot{p}_t,\bm{\theta}) dt + b (X_t;p_t, \bm{\theta} ) dW_t \quad t > 0 \\
X_0 & = x_0
\end{split}
\label{main}
\end{equation}

\begin{itemize}
\item $a(\cdot; p_t, \dot{p}_t,\bm{\theta}):[0,1] \to \R $  a drift function.
\item $b (\cdot;p_t, \bm{\theta} ):[0,1] \to \R$  a  diffusion function.
\item $\bm{\theta}$: a vector of parameters.
\item $p_t$ time-dependent scalar value and $ \dot{p}_t$ is its time derivative at time $t$. (in our case  $p_t$ is a deterministic forecast).
\item $W_t$: Standard Wiener random process in $\R$.
\end{itemize}

@@

\begin{frame}\frametitle{How do we choose an appropriate drift and diffusion functions?}

Let $\textcolor{red}{ \bm{\theta} }= (\textcolor{red}{\theta_0}, \textcolor{red}{\alpha})$
\begin{enumerate}
  \item We want the process to follow the wind forecast, thus we choose a drift term that is mean reverting and tracks the derivative of the deterministic forecast $p_t$, which is an input to our model.
  \begin{equation}
    a(x; p_t,\textcolor{red}{ \bm{\theta} })=  \dot{p_t} - \theta_t(x - p_t)
  \end{equation}
where $\theta_t>0$ is a time-dependent parameter that controls the speed of reversion.
  \item We want a diffusion term that vanishes at the boundaries to prevent the process from escaping the region $[0,1]$.
  \begin{equation}
    b (x; p_t,\textcolor{red}{ \bm{\theta} } )= \sqrt{2 \theta_t \textcolor{red}{\textcolor{red}{\alpha}} x (1-x)}
  \end{equation}
  where $\textcolor{red}{\alpha} >0$ is a constant parameter that controls the path variability.

To further ensure that the process does not escape the region $[0,1]$, the mean reversion parameter has to be selected according to the following rule,
\begin{equation}
\theta_t = \max \left( \textcolor{red}{\theta_0} \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right ),  \quad \textcolor{red}{\theta_0} >0 \label{theta_t}
\end{equation}
\end{enumerate}
\end{frame}

@@

\begin{equation}
\mathcal{L}(\bm{\theta};V) =\prod\limits_{j=1}^M \prod\limits_{i=1}^N \rho ( {V_{j,i+1}|V_{j,i}},\textcolor{red}{ \bm{\theta} })  \rho (V_{j,0})
\label{likelihood}
\end{equation}

@@

\begin{equation}
\begin{split}
\frac{ \partial f }{\partial t } & (y ,t | x , s, \textcolor{red}{\theta_0} , \textcolor{red}{\alpha} )= - \frac{\partial}{ \partial y} ( a( y;  p_t, \dot{p}_t, \textcolor{red}{\theta_0} ) f( y ,t | x , s, \textcolor{red}{\theta_0}  , \textcolor{red}{\alpha} ) ) \\
& + \frac{1}{2} \frac{\partial^2}{ \partial y^2} ( b^2(y; p_t, \textcolor{red}{\theta_0}, \textcolor{red}{\alpha}  )  f(y ,t | x , s,\textcolor{red}{\theta_0} , \textcolor{red}{\alpha} ) ) \quad  t < s\\
\end{split}
\end{equation}
This is a parametric PDE which is \textcolor{red}{computationally expensive to solve} and optimize for every transition.
\end{frame}

\begin{frame}\frametitle{Moment Matching}
We propose a \textcolor{red}{proxy transition density}. We match the moments of our SDE model with that of the proxy density. Using It$\hat{o}$ formula, we arrive at the following iterative ODEs.

\begin{equation}
\begin{split}
\frac{d \E[ V^k_t]}{dt} = - k \theta_t \E [ V^k_t] + \frac{k(k-1)}{2} \E [ V^{k-2}_t  b(V^k_t;\theta_t, \textcolor{red}{\alpha})]
\end{split}
\end{equation}
For $t\in [t_{n-1}, t]$, the first two moments are given by
\begin{equation}
\begin{split}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t \\
\frac{d m_2 (t)}{dt} &=  -2 m_2(t)\theta_t(1+\textcolor{red}{\alpha}) + 2\textcolor{red}{\alpha}\theta_t m_1(t)(1-2p_t) + 2 \textcolor{red}{\alpha}\theta_t p_t (1-p_t)\\
\end{split}
\end{equation}
with initial conditions, $m_1(t_{n-1})= v_{n-1}$ and $m_2(t_{n-1})= v_{n-1}^2$.

@@

Execute the following until an accuracy threshold is met:
\begin{enumerate}
\item \textcolor{red}{initialize}.
\item \textcolor{red}{optimize} the log-likelihood function.
\begin{enumerate}
\item For every evaluation of the log-liklihood function:
\begin{enumerate}
  \item \textcolor{red}{Sample} a mini-batch of transitions randomly with their associated forecast and parameters.
  \item \textcolor{red}{Solve} the ODE system for every transition to obtain the moments.
  \item \textcolor{red}{Match} the resulting moments with the parameters of the chosen proxy distribution for every transition.
\end{enumerate}
\end{enumerate}
\item go to step 1 (i.e. \textcolor{red}{re-initialize} the optimization with the most recent result ).
\end{enumerate}

@@

\begin{frame}\frametitle{Future Wind Power Simulation}
    \begin{figure}
      \includegraphics[width=75mm,scale=1]{simulated/24hr/1099.pdf}
      \includegraphics[width=75mm,scale=1]{simulated/24hr/1178.pdf}
      \caption{We simulate five possible future wind power production paths using the obtained optimal parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha} )=(12,0.3)$ }
    \end{figure}
\end{frame}

\begin{frame}\frametitle{Confidence Bands}
    \begin{figure}
      \includegraphics[width=75mm,scale=1]{confidence_intervals/24hr/31.pdf}
      \includegraphics[width=75mm,scale=1]{confidence_intervals/24hr/820.pdf}
       \caption{We obtain confidence intervals for future wind power production using the obtained optimal parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha} )=(12,0.3)$. Actual production plotted in retrospect. }
    \end{figure}
\end{frame}

\begin{frame}\frametitle{State-Independent Diffusion Formulation}
The model we have demonstrated previously is \textcolor{red}{state-dependent diffusion} formulation, that is the diffusion of the SDE depends on the state.
\begin{equation}
\begin{split}
dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \textcolor{red}{\alpha} (V_t +p_t ) (1-V_t-p_t)} \  dW_t  \\ %\quad t > 0
V_0 & = v_0\\
\end{split}\label{VtSDE}
\end{equation}

Why are we interested in a state-independent diffusion formulation? Because it's more tractable and numerically stable.\\
We apply a \textbf{Lamperti transform} to obtain the following \textcolor{red}{state-independent diffusion} SDE,

\begin{equation}
  \begin{split}
    dZ_t&= \frac{- \theta_t (1+ \sin(Z_t) - 2p_t) + \textcolor{red}{\alpha} \theta_t \sin (Z_t)   }{\cos (Z_t)} \ dt + \sqrt{2 \textcolor{red}{\alpha} \theta_t} dW_t \\
    Z_0&=z_0
  \end{split}
\end{equation}
where $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.
\end{frame}

\begin{frame}\frametitle{Data Skewness after Lamperti Transformation}
As stated before, the state-independent diffusion SDE follows a Lamperti transformed process $Z_t$ given by $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.
\begin{figure}
  \includegraphics[width=50mm,scale=1]{plots/hist_low_lamperti.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_mid_lamperti.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_high_lamperti.pdf}
  \caption{We observe that skewness has been greatly reduced after the Lamperti transformation. This motivates us to use a \textcolor{red}{Gaussian transition} density as a proxy density.}
\end{figure}
\end{frame}

\begin{frame}\frametitle{State-Independent diffusion formulation}
Similarly, we try to obtain a system of ODEs to determine the centered moments of the Lamperti transformed process $V_t$. Due to the non-linearity in the drift, we can only approximate the centered moments by the following ODEs,
\begin{equation}
\begin{split}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t (1-\textcolor{red}{\alpha}) - \theta (1-2 p_t) \\
\frac{d var(t)}{dt} &=  2 var(t) \theta_t (2p_t - 1 ) \tan(m_1 (t)) \sec(m_1 (t)) + \theta_t (\textcolor{red}{\alpha} - 1) \sec^2(m_1 (t))  + 2 \theta_t \textcolor{red}{\alpha}\\
\end{split}
\end{equation}
with initial conditions, $m_1(t_{n-1})= v_{n-1}$ and $var(t_{n-1})= v_{n-1}^2 - v_{n-1}$.\\
\bigskip 
\textcolor{red}{These are not exact ODEs for the centered moments}, however they are accurate enough for small time intervals. 
\end{frame}

\begin{frame}\frametitle{Result Comparison in the Different Spaces}
\begin{table}[]
\centering
\begin{tabular}{|c|c|}
\hline
Formulation   &  parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha})$    \\ \hline
Without Lamperti transform &   $(12,0.3) $   \\ \hline
With Lamperti transform &   $(12,0.29)  $   \\ \hline
\end{tabular}
\caption{We compare the parameters obtained in both the original and Lamperti space. }
\label{tab:model_comparison}
\end{table}
\end{frame}

\begin{frame}\frametitle{Concluding Remarks}
We were able to:
\begin{itemize}
  \item simulate future wind power production based on real data.
  \item obtain an analytical description of the uncertainty of wind power forecasts in the form of an SDE.
  \item develop a forecasting technology agnostic method.
  \item capture skewness of the error process and its dynamics.
\end{itemize}
\end{frame}

@@

\section*{References}
\begin{flushleft}
\begin{itemize}
\item Alhaddad, W. ,  Kebaier, A. , \& Tempone, R. (2019). Stochastic Wind Power Forecasting. In preparation.
\item  M\o ller, J. K., Zugno, M., \& Madsen, H. (2016). Probabilistic Forecasts of Wind Power Generation by Stochastic Differential Equation Models. Journal of Forecasting, 35(3), 189-205. doi:10.1002/for.2367 
\item  Elkantassi, S., Kalligiannaki, E., \& Tempone, R. (2017). Inference And Sensitivity In Stochastic Wind Power Forecast Models. Proceedings of the 2nd International Conference on Uncertainty Quantification in Computational Sciences and Engineering (UNCECOMP 2017). doi:10.7712/120217.5377.16899
\item Giebel, G., Brownsword, R., Kariniotakis, G., Denhard, M., \& Draxl, C. (2011). The State-Of-The-Art in Short- Term Prediction of Wind Power: A Literature Overview, 2nd edition. ANEMOS.plus. https://doi.org/10.11581/DTU:00000017 
\item Chang, W.-Y. (2014) A Literature Review of Wind Forecasting Methods. Journal of Power and Energy Engineering, 2, 161-168. http://dx.doi.org/10.4236/jpee.2014.24023
\end{itemize}
\end{flushleft}
